---
title: "pacu: Precision Agriculture Computational Utilities - Satellite data"
author: "Caio dos Santos & Fernando Miguez"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{pacu: Precision Agriculture Computational Utilities - Satellite data}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
  
# Specific vignettes

1.  [Satellite data](pacu_sat.html)

2.  [Weather data](pacu_weather.html)

3.   [Yield monitor data](pacu_ym.html)

4. [Frequently asked questions](pacu_faq.html)

# Satellite data
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(pacu)
```

## Interacting with the OData API

### Setting up your credentials

The first step is to register your [Copernicus Data Space Ecosystem](https://dataspace.copernicus.eu/) credentials in the R environment using `pa_initialize_dataspace()`.

```{r initializing-data-space, eval = FALSE}
un <- 'my-username'
pw <- 'my-password'
pa_initialize_dataspace(username = un, password = pw)
```

### Browsing the Data Space catalog

Let us define our area of interest. In this example, we are
interested in a field in Ames, Iowa, U.S.

```{r defining-area-of-interest}
extd.dir <- system.file("extdata", package = "pacu")
area.of.interest <- sf::st_read(file.path(extd.dir, 'cobs_a_aoi.shp'), quiet = TRUE)
```

Now we can browse the Data Space catalog for images that
meet our requirements by passing arguments to
`pa_browse_dataspace()`. The function lets us filter
satellite products by date, cloud cover, and collection. In
this example, we are interested in Sentinel-2 images from
2023 with at most 50% cloud cover.

```{r browsing-data-space, eval = FALSE}
available.images <- pa_browse_dataspace(aoi = area.of.interest,
                                        max.cloud.cover = 50,
                                        start.date = '2023-01-01',
                                        end.date = '2023-12-31',
                                        collection.name = 'SENTINEL-2')

```

Each entry in the data frame returned by the function is a
satellite image available to be downloaded from Data Space.
In this case, there are 69 images that meet our
requirements.

```{r reading-browse-ds, include=FALSE}
available.images <- readRDS(file.path(extd.dir, 'ds-browse-object.rds'))
```

```{r inspecting-browse-ds}
available.images
```

Using the "summary" function, we can extract the number of
available images for every month in the data set.
```{r summary-browse-ds}
summary(available.images)
```


### Downloading satellite images from Data Space

The next step is to download the images. If we supply the
`available.images` object to `pa_download_dataspace()`, the
function will download all 69 available images. For this
demonstration, we can focus only
on two images. Let's pick the 32nd and 33rd images because
these were captured in mid-July, so there should be maize or
soybeans in the field.

An important consideration is that these images can occupy a
considerable amount of storage, roughly 500 MB each. That
means downloading all 69 images would require about 35 GB.
If you provide an area of interest (`aoi`), the function
will crop the downloaded imagery to the relevant extent,
which helps reduce storage needs when working with many
images.

```{r downloading-an-image, eval=FALSE}
out.dir <- tempdir()
available.images <- available.images[32:33, ]
pa_download_dataspace(x = available.images,
                      dir.path = out.dir,
                      aoi = area.of.interest,
                      verbose = FALSE)
```

```{r out-equal-ext, include=FALSE}
out.dir <- extd.dir
```

### Visualizing downloaded images

Now that we have downloaded the data, we might be interested
in looking at RGB images.

```{r rgb-img, fig.width=7, fig.height=5}
s2.files <- list.files(out.dir, '\\.zip', full.names = TRUE)
rgb.img <- pa_get_rgb(s2.files,
                      aoi = area.of.interest,
                      verbose = FALSE)

pa_plot(rgb.img)
```

Similarly, we might be interested in NDVI or NDRE.
```{r, ndvi-img,  fig.width=7, fig.height=5}
ndvi.img <- pa_compute_vi(s2.files,
                          vi = 'ndvi', 
                          aoi = area.of.interest,
                          verbose = FALSE)

ndre.img <- pa_compute_vi(s2.files, 
                          vi = 'ndre',
                          aoi = area.of.interest,
                          verbose = FALSE)

pa_plot(ndvi.img, main = 'NDVI')
pa_plot(ndre.img, main = 'NDRE')
```

## Area-level summary

In many precision agriculture applications, we are
interested in a summary value for each experimental unit to
evaluate the effect of a treatment. This statistic is often
the mean or median of the pixels within a polygon
representing a research plot or another area of interest.
To illustrate this in *pacu*, we can split the area of
interest into four smaller polygons and then compute and
visualize summary statistics.

```{r  ndvi-img-summary,  fig.width=7, fig.height=5}
split.aoi <- st_make_grid(area.of.interest, n = c(4, 1))
split.aoi <- st_as_sf(split.aoi)
split.aoi <- st_transform(split.aoi, st_crs(ndvi.img))
ndvi.mean <- summary(ndvi.img, by = split.aoi, fun = mean)
pa_plot(ndvi.mean)
```

## Interacting with the Copernicus Statistical API

The Data Space Statistics API allows users to request
area-level statistics derived from satellite imagery
without downloading the imagery itself. The API uses
OAuth2.0 authentication with a client ID and client secret.
To use these functions, you will need to [register an OAuth client](https://documentation.dataspace.copernicus.eu/APIs/SentinelHub/Overview/Authentication.html).

### Setting up your credentials
 
```{r registering-oauth, eval = FALSE}
cid <- 'my-client-id'
cs <- 'my-client-secret'
pa_initialize_oauth(client_id = cid, client_secret = cs)
```

### Requesting areal statistics

Now that your OAuth client credentials have been registered,
you can request vegetation index statistics using the
Statistics API. Let us request NDVI data for our area of
interest.

```{r requesting-ndvi-statistics, eval = FALSE}
ndvi.statistics <- pa_get_vi_stats(aoi = area.of.interest,
                                   start.date = '2022-01-01',
                                   end.date = '2022-12-31',
                                   vegetation.index = 'ndvi',
                                   agg.time = 'P1D')
```

```{r reading-ndvi-statistics, include = FALSE}
ndvi.statistics <- readRDS(file.path(extd.dir, 'example-ndvi-stats.rds'))
```

The `pa_plot()` function helps visualize the results in a
spatial context. Because we requested NDVI statistics for a
single polygon, each facet representing a date is expected
to contain only one color.

```{r plotting-ndvi-statistics-paplot, fig.width=7, fig.height=5}
pa_plot(ndvi.statistics)
```


If we are more interested in a time-series view, we can
specify `plot.type`.
```{r vis-ndvi-stats, fig.width=7, fig.height = 5}
pa_plot(ndvi.statistics, 
        plot.type = 'timeseries')
```



We might also want to compare index values across different
regions of the field. Let us split the previous field into
four parts and retrieve NDVI separately for each section.


**Note:** using the `by.feature` argument sends multiple
requests to the Statistical API, so you should be mindful
of your quota. Quotas and limitations are described
[here](https://documentation.dataspace.copernicus.eu/Quotas.html).

```{r requesting-ndre-statistics, eval = FALSE}
ndvi.statistics.2 <- pa_get_vi_stats(aoi = split.aoi,
                                     start.date = '2022-01-01',
                                     end.date = '2022-05-01',
                                     vegetation.index = 'ndvi',
                                     agg.time = 'P1D',
                                     by.feature = TRUE)
```

```{r reading-ndre-statistics, include = FALSE}
ndvi.statistics.2 <- readRDS(file.path(extd.dir, 'example-ndvi-stats-2.rds'))
```


Now we can visualize the different parts of the field in a
spatial and temporal context.
```{r plotting-ndvi-statistics-multiple, fig.width=7, fig.height=5}
pa_plot(ndvi.statistics.2)
```


Similarly, we can examine the same data in a time-series
format. In this case, specifying `by = "id"` plots a
different series for each polygon.
```{r plotting-ndvi-statistics-multiple-timeseries, fig.width=7, fig.height=5}
pa_plot(ndvi.statistics.2, 
        plot.type = 'timeseries',
        by = c('id'),
        legend.outside = TRUE)
```



