---
title: "pacu: Precision Agriculture Computational Utilities - FAQ"
author: "Caio dos Santos & Fernando Miguez"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{pacu: Precision Agriculture Computational Utilities - FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Specific vignettes

1.  [Satellite data](pacu_sat.html)

2.  [Weather data](pacu_weather.html)

3.   [Yield monitor data](pacu_ym.html)

4. [Frequently asked questions](pacu_faq.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacu)
```

# FAQ

1.  The *pa_yield()* function expects an *sf* object, but my data is in `.csv` format. How should I proceed?

    Data stored in CSV format can be read into R as a data frame and then converted to an *sf* object. See the example below.
    ```{r, message = F}
    x <- 40:45
    y <- -90:-95
    yield <- 100:105
    dat <- data.frame(x = x, 
                      y = y,
                      yield = yield)
    class(dat) ## data.frame
    
    dat <- sf::st_as_sf(dat,
                 coords = c('x', 'y')) 
    class(dat) ## sf and data.frame
    
    ```

1.  How do I silence warnings and messages within the package?

    To silence warnings and messages, use `pacu_options()`.
    ```{r}
    pacu_options(suppress.warnings = TRUE,
                 suppress.messages = TRUE)
    ```
    
    